---
title: "6. Data tidying"
author: "SHH"
date: "2023-07-11"
output: html_document
---

```{r}
library(tidyverse)
```

# 1. Tidy data
There are three interrelated rules that make a dataset tidy:

1. Each variable is a column
2. Each observation is a row 
3. each value is a cell 

Two main advantages of tidy data: 

1. Consistent way of storing data: easier to learn the tools work with it
2. There's a general advantage to placing variables in columns because it allows R's vectorized nature to shine 

All the packages in the tidyverse are designed to work with tidy data.

<br>

***

<br>

# 2. Lengthing data 
Unfortunately, most real data is untidy. There are two main reasons:

1. Data is often organized to facilitate some goal other than analysis(e.g. data to be structured to make data entry easy)
2. Most people are not familiar with the principle of tidy data

Tidying data steps:

1. figure out what the underlying variables and observations are
2. pivot data into a tidy form

<br>

### Data in column names: pivot_longer()
The `billboard` dataset records the billboard rank in the year 2000. Here, the column names are one variable and the cell values are another.

`pivot_longer` key arguments:

- `cols`: specify which columns need to be pivoted 
- `names_to`: name of the variable stored in the column names
- `values_to`: name of the variable storing in the cell values
- `values_drop_na`: drop the na value

```{r}
billboard
```

```{r}
billboard |> 
  pivot_longer(
    cols = starts_with('wk'),
    names_to = 'week',
    values_to = 'rank'
  )
```

```{r}
billboard |> 
  pivot_longer(
    cols = starts_with('wk'),
    names_to = 'week',
    values_to = 'rank',
    values_drop_na = TRUE
  )
```

This data is now tidy, but we can make future computation a bit easier by converting values of `week` fro m character to numbers using `mutate()` and `readr::parse_number()`.
`parse_number()` is a handy function that will extract the first number from a string, ignoring all other text.
```{r}
billboard_longer <- billboard |> 
  pivot_longer(
    cols = starts_with('wk'),
    names_to = 'week',
    values_to = 'rank',
    values_drop_na = TRUE
  ) |> 
  mutate(week = parse_number(week))
```
```{r}
billboard_longer
```

```{r}
# Very few songs stay in the top 100 for more than 20 weeks
billboard_longer |> 
  ggplot(aes(x = week, y = rank, group = track)) +
  geom_line(alpha = 0.25) +
  scale_y_reverse()
```

<br>

### How deos pivoting work?
```{r}
df <- tribble(
  ~id, ~bp1, ~bp2,
  'A', 100, 120,
  'B', 140, 115,
  'C', 120, 125
)
df
```

```{r}
df |> 
  pivot_longer(
    cols = bp1:bp2,
    names_to = 'measurement',
    values_to = 'value'
  )
```

1. The values in column that was already a variable in the original dataset(id) need to be repeated
2. The column names become values in a new variable
3. The cell values also become values in a new variable

<br>

### Many variables in column names 

- sp/rel/ep: method of diagnosis
- m/f: gender
- 014/1524/2534/3544/.../65: age range 

So in this case we have six pieces of information: `country`, `year`, `method of diagnosis`, `gender`, `age range`, `count`

```{r}
who2
```

```{r}
who2 |> 
  pivot_longer(
    cols = !(country:year),
    names_to = c('diagnosis', 'gender', 'age'),
    names_sep = '_',
    values_to = 'count'
  )
```

